Well-Designed Environmental Markets Enable Large-Scale Marine Conservation

### 2019-10-19 docs/Markets_Conservation_revised.tex
Updated code and docs

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
While it is commonly agreed that marine conservation should expand dramatically around the world, most countries have yet to implement large-scale Marine Protected Areas in their own waters. When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While this may be more-than-made-up-for through spillover of protected fish, these benefits typically accrue to other countries or the high seas. We show that careful design of cross-country rights-based approaches to fisheries can actually incentivize, rather than relegate, large scale marine conservation. By combining a model of cross-country trading of rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that two market design features are pivotal to encouraging such conservation. First, only when trading of fishing rights across countries is allowed, can a conservation-minded country capture the economic spillover benefits of their conservation actions. Second, the allocation of rights to countries must not be watered-down over time by a country's conservation efforts; for example, a country's annual allocation of rights should depend on the size of the fish stock in their waters, not on how hard it has fished. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

Recognizing the need to protect marine biodiversity and ecosystem services, various international bodies have committed to dramatically expand marine protection around the world by making 30\% of the oceans off-limits to fishing \cite{oleary_2016,dinerstein_2019}. While most marine protected areas (MPAs) implemented to date have been small, it is widely-recognized that to meet these goals, very large MPAs must also be part of the strategy. But would any country rationally close 30\%, 80\% or even 100\% of its waters to fishing if this means losing all fishing revenue from within the closed area? At first glance, the answer is probably ``no''. Losing this important source of revenue would cripple many fishing-dependent economies, particularly the Pacific Island nations who have been suggested as viable candidates for large closures \cite{mcleod_2019}. On the other hand, the closure may generate substantial ``spill-over'' of fish and other benefits into adjacent waters which could, in principle, offset these losses. The problem with this argument is that for very large MPAs, these spillover benefits would accrue to other nations, with no obvious mechanism for the conserving country to recoup them. A viable solution may lie in international fishing effort markets, where nations trade the right to fish with each other. The design and effectivness of fishing effort markets for fisheries management have been explore before \cite{havice_2010,havice_2013}, but little attention has been given to the role that markets can play in conservation. Here, we will show how fishing effort markets can be designed (if new) or modified (if already existing) to incentivize implementation of large-scale marine protected areas.

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific island nations that collectively manages tuna purse seine fishing in its members' waters \cite{havice_2013,aqorau_2018}. These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by December 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how the market can be designed to incentivize conservation.

Not all market-based approaches to environmental management are created equal. A pervasive finding across a range of natural resources is that features of markets, such as the allocation of rights, can have implications for the market's functioning \cite{libecap_1989}. In the context of fishing effort markets, we find that two market design features are pivotal in determining the incentives for large-scale marine conservation: trading and allocation rules. But why would these design features of a fisheries market affect the incentives for conservation? Consider the incentives for a country to close 100\% of its waters. Such a closure would surely benefit other countries through the spillover of fish from the protected area to the waters of neighboring countries. If the conserving country could trade the rights to catch the fish that spill-over, other countries would likely buy them, offsetting the costs of not fishing its own waters. But if the conserving country was not allowed to lease these rights, then it would lose all of its fishing revenue. The rules for how fishing rights are allocated across countries are equally important. Suppose rights are allocated each year based on the previous year's fishing effort: the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior. In the following section, we will show how trading and allocation rules can shape the incentives that may facilitate or hinder large-scale conservation. We will then make use of vessel-tracking data to analyze a real-world case of large-scale marine conservation under fishing effort markets.

@@

We have shown that trading significantly reduces the costs of conservation. However, a new question arises. How should rights be re-allocated every year once a country starts closing its waters to fishing? Customarily, the allocation of fishing effort rights is a formula that combines historical fishing effort and biomass in a country's waters \cite{havice_2013}. We test a range of allocation rules that weight effort ($\alpha$) and biomass ($1 - \alpha$) differently as the basis for ongoing rights allocation. We simulate a fishery operating with closures for 50 years and compare the resulting revenues to a fishery without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in long-term losses to the conserving country of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, a biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in revenue losses as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation. This result implies that if allocation is based purely on the biological features of a stock (\emph{e.g.} the biomass within a nation's waters), and not on fishing effort, the incentives for conservation are sustained through time.

@@

A large-scale MPA was recently implemented in PNA waters, providing the ideal empirical setting to test our predictions. In January 2015, Kiribati closed 11.5\% of its EEZ, effectively displacing all fishing effort within its boundaries\cite{mccauley_2016,mcdermott_2018}. By protecting important tuna spawning habitat, PIPA may provide important recruitment and biomass benefits to adjacent waters \cite{hernndez_2019}. We combine vessel-tracking data \cite{kroodsma_2018} and country-level license revenue data \cite{ffa_2017} to quantify the displacement of fishing effort and the likely costs of conservation. Of the 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation and 28 ``non-displaced vessels'' never fished in PIPA waters but fished within Kiribati's EEZ. The remaining 221 vessels were not continuously observed before and after but are included in our analyses, and we refer to these as ``other vessels''.

Consistent with our models' prediction, PIPA caused vessels to relocate largely outside of Kiribati, and into other PNA countries' waters (Figs. S5 - S7). Indeed, displaced vessels fished 922 fewer days (a 10\% reduction) in Kiribati, but spent just 48 fewer days in PNA waters (Fig. \ref{fig:empirical}).  The reduction in effort in Kiribati and constant effort at the PNA-level suggest that trading facilitated redistribution of effort within PNA waters, just as the model predicted. On the other hand, non-displaced and other vessels spent 1,621 and 2,034 additional days in Kiribati in 2015, relative to 2014. The same pattern is observed at the PNA level, with non-displaced and other vessels spending an additional 3,789 and 3,691 vessel-days, respectively. By 2018, we observe a net decrease of vessel-days within Kiribati, from 12,282 in 2014 to just 7,542 in 2018, with displaced vessels driving the decrease (Fig. \ref{fig:empirical}A).

@@

Our findings may help inform management and implementation of existing and upcoming MPAs in the PNA. In December 2020, Palau will close nearly 80\% of its EEZ to commercial fishing to create the Palau National Marine Sanctuary (PNMS): the 14\textsuperscript{th} largest protected area in the world (Fig. \ref{fig:PNA_map}). Vessel-tracking data (2012 to 2018) shows that the proposed PNMS would displace 65.6 $\pm$ 0.08\% ($\pm$ 1SD) of longline vessel-days (non-tradable) and 82.2 $\pm$ 0.08\% ($\pm$ 1SD) of purse seine vessel-days (tradable; Fig. S10-S11).

While trading would allow Palau, Kiribati, and other PNA members to reduce revenue losses from large-scale conservation, our model suggests that they must also adopt a biomass-based allocation rule to ensure long-term financial security in the presence of large-scale conservation. The Parties to the Nauru Agreement have shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and allocating rights based on biomass, they may become pioneers in effective large-scale marine conservation in a market-based setting.

In our study, only one of the countries considers the implementation of a protected area. However, an interesting extension would be cooperative conservation. In that setting, a group of countries could coordinate on an optimized large-scale strategical closure, perhaps using vessel-day trading as a means of compensation, in a matter similar to the model above. Future research could explore the role of heterogeneous costs and benefits between countries, and the way in which these can shape conservation outcomes in a market-based setting.

@@

We model a 10 country discrete-time meta-population system, where Country 1 is considering a spatial closure. Countries 1 - 9 operate under a vessel-day scheme, and Country 10 represents the high seas and other areas not managed under a VDS. The stock of fish in each country is relatively stationary within a single fishing season, but  growth from escapement redistributes across all countries annually. The price of fish is $p$, and catchability is given by $q$. These parameters are held constant across countries.

@@

In the absence of a reserve, the revenue for vessels in country $i$ is given by $pqE_iX_i$, where $E_i$ and $X_i$ are effort (vessel-days) and stock size in country $i$ at the beginning of a period. The cost of fishing in country $i$ is given by $cE_i^\beta$, where $\beta = 1.3$ matches commonly-used cost functions.

Country 1 considers a spatial closure by implementing a reserve as a fraction $R$ of the total country ($R \in[0,1)$). Fish move within a country based on $\theta$, where $\theta = 0$ implies no movement within the country, and $\theta = 1$ implies that fish move so much that they can be caught from anywhere within the country. In this country, revenues are given by $pqE_1X_1(\theta + (1 - \theta)(1 - R))$. The parameterization of movement and reserve size imply that profit from fishing Country 1 is given by:

@@

@@

Annual vessel-days are often allocated based on a combination of historical within-country effort and biomass. In the PNA, for example, 60\% of the allocation is calculated based on EEZ effort over the last seven years and 40\% is calculated based on the 10-year average of each country’s share of estimated skipjack and yellowfin biomass within its EEZ (This is explained in more detail in Article 12.5 of the 2012 Amendment to the Palau Agreement and in \cite{Hagrannsoknir2014}). Trading vessel-days to other countries would imply that historical within-country effort declines through time. The allocated days to a country with a full spatial closure would eventually be reduced to just the 40\% based on biomass.

@@

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcast messages can be received by satellites and land-based antennas. We use AIS data provided by Global Fishing Watch \cite{kroodsma_2018} to track 313 tuna purse seiners that fished within the PNA. Of the 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation, 28 ``non-displaced vessels'' never fished in PIPA waters, and  221 ``other vessels'' were not continuously observed before and after. Together, these vessels represent more than 20 million geo-referenced positions.

We use these data to calculate the number of vessel-days that the 313 purse seiners spent in each PNA country and in PNA waters as a whole. The vessel-day equivalent of a day of fishing depends on vessel size, a measure used to control for effort creep. A day of activity by vessels smaller than 50 meters long counts as half a vessel-day, a day of activity by vessels 50 - 80 meters counts as one day, and a day of activity by vessels larger than 80 meters counts as 1.5 vessel-days. Vessel length is an observable characteristic in our dataset, and therefore we make the pertinent corrections when calculating vessel days.

Shapefiles of Exclusive Economic Zones were obtained through Marine Ecoregions of The World, we use World EEZ v10 (2018-02-21) available for download at: http://www.marineregions.org. Shapefiles for Marine Protected Areas (Phoenix Islands Protected Area and Palau National Marine Sanctuary) come from the World Database of Protected Areas, and were downloaded in March 2019 from: https://www.protectedplanet.net. All analyses were performed in R version 3.5.3 \cite{rcore_2018}.

@@

We obtained information on revenues from the Pacific Islands Forum Fisheries Agency \emph{Tuna Development Indicators 2016} report.  Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA\cite{ffa_2017}) where annual revenues from license fees (for VDS and other access programs) are reported for each country (2008 - 2016; Fig. \ref{fig:empirical}C; Fig. S8 - S9). For countries in the PNA, these revenues show a combination of vessel-day license fees as well as joint-venture operations.

@@

Juan Carlos Villaseñor-Derbez was supported by UCMexus-CONACyT (CVU: 669403) and the Latin American Fisheries Fellowship Program. John Lynham was supported by the Conservation Strategy Fund.

@@

% PNA model output for costs
\begin{figure}[htbp]
\centering
\includegraphics{img/PNA_model.pdf}
\caption{\label{fig:PNA_model}Cost of spatial closures in a vessel-day fishery. Each line represents a possible value of within-country stock movement ($\theta$; line color). The revenue losses to Country 1 (vertical axis) are relative to a fishery with no spatial closures, and are shown as a function of reserve size ($R$; horizontal axis) and movement (color). Costs are shown for Country 1 when there is no trading (A) and when trading is allowed (B). Costs avoided by trading are shown in (C). Dashed red line in (A) is a 1:1 line.}
\end{figure}

% Costs of different allocation rules
\begin{figure}[htbp]
\centering
\includegraphics{img/allocation_cost_plot.pdf}
\caption{\label{fig:allocation_cost_plot}Costs of a spatial closure for Country 1 under different allocation rules. Each line represents the revenue losses for a combination of allocation rules ($\alpha$; line type) and movement ($\theta$; color) for different reserve sizes ($R$; horizontal axis). An effort-based allocation and low movement values result in the highest costs. Cost can be minimized for all movement values if allocation is based on biomass within each country’s waters.}
\end{figure}

% Effort redistribution bars for PNA and KIR
\begin{figure}[htbp]
\centering
\includegraphics{img/empirical.pdf}
\caption{\label{fig:empirical}Effort displacement and license revenues. Panels (A) and (B) show AIS-derived annual vessel-days for Kiribati and for all Parties to the Nauru Agreement (PNA) by 313 tuna purse seine vessels. Annual effort is broken down by displaced (n = 64), non-displaced (n = 28), and other vessels (n = 221). The dashed horizontal lines represent the total allowable effort in Kiribati (11,000 days \cite{yeeting2018stabilising}) and the PNA (45,000 days). Panel (C) shows annual revenue from fishing license fees by country and year (2008 - 2016) and panel (D) shows the correspondence between FFA-reported revenues and AIS-derived vessel-day observations (2012 - 2016). The dashed line represents line of best fit.}
\end{figure}

>>>>

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
While it is commonly agreed that marine conservation should expand dramatically around the world, most countries have yet to implement large-scale Marine Protected Areas in their own waters. When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While this may be more-than-made-up-for through spillover, these benefits typically accrue to other countries or the high seas. We show that careful design of cross-country rights-based approaches to fisheries can actually incentivize, rather than relegate, large scale marine conservation. By combining a model of cross-country trading of rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that two market design features are pivotal to encouraging such conservation. First, only when trading of fishing rights across countries is allowed, can a conservation-minded country capture the economic spillover benefits of their conservation actions. Second, the allocation of rights to countries must not be watered-down over time by a country's conservation efforts; for example, a country's annual allocation of rights should depend on the size of the fish stock in their waters, not on how hard it has fished. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

Recognizing the need to protect marine biodiversity and ecosystem services, various international bodies have committed to dramatically expand marine protection around the world by protecting 30\% of the oceans \cite{oleary_2016,dinerstein_2019}, and that large portions of these will be entirely no-take \cite{sala_2017,sala_2018}. While most marine protected areas (MPAs) implemented to date have been small, it is widely-recognized that to meet these goals, very large MPAs must also be part of the strategy. But would any country rationally close 30\%, 80\% or even 100\% of its waters to fishing if this means losing all fishing revenue from within the closed area? At first glance, the answer is probably ``no''. Losing this important source of revenue would cripple many fishing-dependent economies, particularly the Pacific Island nations who are viewed as viable candidates for large closures \cite{mcleod_2019}. On the other hand, the closure may generate substantial ``spillover'' of larvae and adult fish \cite{ramesh_2019,hernndez_2019} and other benefits into adjacent waters which could, in principle, offset these losses. The problem with this argument is that for very large MPAs, these spillover benefits could accrue to other nations and the high seas \cite{agardy_2018}, with no obvious mechanism for the conserving country to recoup them. A viable solution may lie in international fishing effort markets, where nations trade the right to fish with each other. The design and effectivness of fishing effort markets for fisheries management have been explored before \cite{havice_2010,havice_2013}, but little attention has been given to the role that markets can play in conservation. Here, we will show how fishing effort markets can be designed (if new) or modified (if already existing) to incentivize implementation of large-scale marine protected areas.

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific island nations that collectively manages tuna purse seine fishing in its members' waters (Fig. \ref{fig:PNA_map}) \cite{havice_2013,aqorau_2018}. These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by December 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how the market can be designed to incentivize conservation.

Not all market-based approaches to environmental management are created equal. A pervasive finding across a range of natural resources is that features of markets, such as the allocation of rights, can have implications for the market's functioning \cite{libecap_1989}. In the context of fishing effort markets, we find that two market design features are pivotal in determining the incentives for large-scale marine conservation: trading and allocation rules. But why would these design features of a fisheries market affect the incentives for conservation? Consider the incentives for a country to close 100\% of its waters. Such a closure would surely benefit other countries through the spillover of fish from the protected area to the waters of neighboring countries. If the conserving country could trade the rights to catch the fish that spillover, other countries would likely buy them, offsetting the costs of not fishing its own waters. But if the conserving country was not allowed to lease these rights, then it would lose all of its fishing revenue. The rules for how fishing rights are allocated across countries are equally important. Suppose rights are allocated each year based on the previous year's fishing effort: the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior. In the following section, we will show how trading and allocation rules can shape the incentives that may facilitate or hinder large-scale conservation. We will then make use of vessel-tracking data to analyze a real-world case of large-scale marine conservation under fishing effort markets.

@@

We have shown that trading significantly reduces the costs of conservation. However, a new question arises. How should rights be re-allocated every year once a country starts closing its waters to fishing? Customarily, the allocation of fishing effort rights is a formula that combines historical fishing effort and biomass in a country's waters \cite{havice_2013}. We test a range of allocation rules that weigh effort ($\alpha$) and biomass ($1 - \alpha$) differently as the basis for ongoing rights allocation. We simulate a fishery operating with closures for 50 years and compare the resulting revenues to a fishery without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in long-term losses to the conserving country of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, a biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in revenue losses as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation. This result implies that if allocation is based purely on the biological features of a stock (\emph{e.g.} the biomass within a nation's waters), and not on fishing effort, the incentives for conservation are sustained through time.

@@

A large-scale MPA was recently implemented in PNA waters, providing the ideal empirical setting to test our predictions. In January 2015, Kiribati closed 11.5\% of its EEZ to implement the Phoenix Island Protected Area (PIPA), effectively displacing all fishing effort within its boundaries\cite{mccauley_2016,mcdermott_2018}. By protecting important tuna spawning habitat, PIPA may provide important recruitment and biomass benefits to adjacent waters \cite{hernndez_2019}. We combine vessel-tracking data \cite{kroodsma_2018} and country-level license revenue data reported by the Pacific Islands Forum Fisheries Agency (FFA) \cite{ffa_2017} to quantify the displacement of vessel-days and the likely costs of conservation. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation and 254 ``non-displaced vessels'' never fished in PIPA waters but fished within the PNA waters. We use the tracking data to calculate vessel-days in the same way that the PNA does (see Methods for details) and the redistribution of fishing events before and after the implementation of PIPA.

Consistent with our models' prediction, after PIPA was implemented, displaced vessels relocated largely outside of Kiribati, and into other PNA countries' waters (Fig \ref{fig:fishing_raster_diff} and Figs. S6). Indeed, from 2014 to 2015, displaced vessels spent 2,115 fewer vessel-days (a 25\% reduction) in Kiribati, and 2,298 fewer vessel-days in PNA waters (a 17 \% reduction; Fig. \ref{fig:empirical}). On the other hand, non-displaced vessels spent 4,656 additional days in Kiribati during 2015, and an additional 9,598 vessel-days in PNA waters. By 2018, we observe a net decrease of vessel-days within Kiribati, from 12,671 in 2014 to just 7,677 in 2018, with displaced vessels driving the decrease (Fig. \ref{fig:empirical}A). However, aggregate effort at the PNA-level remains relatively constant and we do not observe a ``fishing the line'' effect (Fig S7). The reduction in effort to Kiribati and constant effort at the PNA-level suggest that trading facilitated redistribution of effort within PNA waters, just as the model predicted.

The decrease of vessel-days in Kiribati could be alternatively (or jointly) explained by changes in oceanographic conditions that drive distribution of target species and fishing vessels. For example, as El Niño events develop, tuna species are known to shift longitudinally across PNA waters, causing vessels to redistribute \cite{aqorau_2018,hanich2018unraveling}. However, the total decrease that we observe for Kiribati in Fig. \ref{fig:empirical} is driven by the relocation of vessels that once fished within PIPA, not the entire fleet. One could argue that differences in vessel characteristics between displaced and non-displaced vessels may influence their ability to redistribute or take advantage of different oceanographic conditions in Kiribati. For example, on average, displaced vessels have smaller crew sizes, more engine power, and are larger than non-displaced vessels (Table S2, Fig. S5). However, displaced vessels gradually reduce their time spent in PNA waters too.

@@

Our findings may help inform management and implementation of existing and upcoming MPAs in the PNA. In January 2020, Palau will close nearly 80\% of its EEZ to commercial fishing to create the Palau National Marine Sanctuary (PNMS): the 14\textsuperscript{th} largest protected area in the world (Fig. \ref{fig:PNA_map}). Vessel-tracking data (2012 to 2018) shows that the proposed PNMS boundaries contain 65.6 $\pm$ 0.08\% ($\pm$ 1SD) of longline vessel-days (non-tradable) and 82.2 $\pm$ 0.08\% ($\pm$ 1SD) of purse seine vessel-days (tradable; Fig. S10-S11).

%%%% PLEASE TAKE A LOOK AT THE SECOND SENTENCE OF THIS PARAGRAPH %%%%
While trading would allow Palau, Kiribati, and other PNA members to reduce revenue losses from large-scale conservation, our model indicates that moving from a 40\% biomass-based allocation rule to a 100\% biomass-based rule would ensure long-term financial security in the presence of large-scale MPAs, and further incentivize conservation within the PNA. However, it is worth mentioning that we did not empirically test this prediction, as this would imply modifying the allocation rule in the real world. The Parties to the Nauru Agreement have shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and allocating rights based on biomass, they may become pioneers in effective large-scale marine conservation in a market-based setting.

There are, however, a series of important considerations to take into account. Firstly, in our study only one of the countries considers the implementation of a protected area. However, an interesting extension would be that of cooperative conservation. In that setting, a group of countries could coordinate on an optimized, large-scale strategic closure -perhaps using vessel-day trading as a means of compensation- in a manner similar to the model above. Future research could explore the role of heterogeneous costs and benefits between countries, and the way in which these can shape conservation outcomes in a market-based setting. A second consideration is that of the role that the high seas play: environmental markets require secure property rights, which are lacking in the high seas \cite{crespo_2019}. Any spillover from into the high seas could potentially be eroded by the prevailing open access conditions. Therefore, a conservation-minded nation has no mechanism to capture the benefits provided to the high seas, highligting the importance of empowering high seas governance and transboundary cooperation \cite{aqorau_2018,crespo_2019}.

@@

We model a 10 country discrete-time meta-population system, where Country 1 considers a spatial closure. Countries 1 - 9 operate under a vessel-day scheme, and Country 10 represents the high seas and other areas not managed under a VDS. The stock of fish in each country is relatively stationary within a single fishing season, but  growth from escapement redistributes across all countries annually. The price of fish is $p$, and catchability is given by $q$. These parameters are held constant across countries.

@@

In the absence of a reserve, the revenue for vessels in country $i$ is given by $pqE_iX_i$, where $E_i$ and $X_i$ are effort (vessel-days) and stock size in country $i$ at the beginning of a period. The cost of fishing in country $i$ is given by $cE_i^\beta$, where $\beta = 1.3$ matches commonly-used cost functions that assume increasing units of effort are increasingly costly to apply \cite{costello_2016}.

Country 1 considers a spatial closure by implementing a reserve as a fraction $R$ of the total country ($R \in[0,1)$). Fish move within a country based on $\theta$, where $\theta = 0$ implies no movement within the country, and $\theta = 1$ implies that fish move so much that they can be caught from anywhere within the country (see the \emph{Notes on within-country fish movement} section below). In this country, revenues are given by $pqE_1X_1(\theta + (1 - \theta)(1 - R))$. The parameterization of movement and reserve size imply that profit from fishing Country 1 is given by:

@@

\subsubsection{Notes on within-country fish movement}

In our model parameterization, the proportion of biomass ``available'' for harvest is given by  $\Omega_1 = (\theta + (1 - \theta)(1 - R))$. This implies that for a given country with stock size $X_i$, the total biomass available for harvest will be given by $\Omega X_i$. Consider the case of a sesile fish with $\theta = 0$. If the country were to close 50\% of its waters to fishing ($R = 0.5$), only 50\% of the stock would be available for harvest (\emph{i.e.} $0 + (1 - 0) * (1 - 0.5) = 0.5$). Now, consider the same closure is applied with a stock with high mobility, so $\theta = 0.9$. In this case, despite the closure, fish constantly swim between the reserve and fishing grounds making 95\% of biomass available for harvest (\emph{i.e.} $0.9 + (1 - 0.9) * (1 - 0.5) = 0.95$). As derived in the bioeconomic model above, a vessel's willingness to pay to fish in a given patch will be determined by the amount of biomass available for harvest. Within-patch stock movement therefore plays an important role in determinig a vessel's willingness to pay in the remaining open waters: a vessel will be willing to pay more to fish in waters where 95\% of biomass is available for harvest, than when only 50\% of biomass is available for harvest.

@@

Annual vessel-days are often allocated based on a combination of historical within-country effort and biomass. In the PNA, 60\% of the allocation is calculated based on EEZ effort over the last seven years and 40\% is calculated based on the 10-year average of each country’s share of estimated biomass (of skipjack and yellowfin tuna) within its EEZ (This is explained in more detail in Article 12.5 of the 2012 Amendment to the Palau Agreement and in \cite{Hagrannsoknir2014}). Trading vessel-days to other countries would imply that historical within-country effort declines through time. The allocated days to a country with a full spatial closure would eventually be reduced to just the 40\% based on biomass.

@@

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcast messages can be received by satellites and land-based antennas. We use AIS data provided by Global Fishing Watch \cite{kroodsma_2018} to track 318 tuna purse seiners that fished within the PNA. For every georeferenced position, we observe the time spent (defined as the time since the last position), and whether the vessel was activly fishing versus only transiting. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation, the remianing 254 ``non-displaced vessels'' never fished in PIPA waters. Our dataset contains more than 37 million geo-referenced positions for these 318 tuna purse seiners. We use these data to calculate vessel-days (the metric used by the PNA), and to track the spatial redistribution of displaced vessels. A comparision of vessel characteristics between displaced and non-displaced vessels is presented in Table S2 and Figure S5.

We use these data to calculate the number of vessel-days that the 318 purse seiners spent in each PNA country and in PNA waters as a whole (Fig. \ref{fig:empirical}). The vessel-day equivalent of a day of fishing depends on vessel size, a measure used to control for effort creep. A day of activity by vessels smaller than 50 meters long counts as half a vessel-day, a day of activity by vessels 50 - 80 meters counts as one day, and a day of activity by vessels larger than 80 meters counts as 1.5 vessel-days. Vessel length is an observable characteristic in our dataset, and therefore vessel-days calculated in our analyses correspond to the PNA definition of vessel-days.

We can then compare the footprint of displaced and non-displaced vessels before and after the implementation of PIPA to better understand the effort redistribution. Non-displaced vessels serve as a plausible control group that was not subject to a spatial closure but might have redistributed in response to changing environmental conditions, such as El Niño \cite{hanich2018unraveling,aqorau_2018}.

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell, resulting in 14 (7 years for two groups) gridded datasets of fishing hours. For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group (Fig. \ref{fig:fishing_raster_diff}A-B). The redistribution by non-displaced vessels (Fig. \ref{fig:fishing_raster_diff}B) therefore provides a baseline of redistribution. We then compare the changes of displaced vessels to those by non-displaced vessels (Fig. \ref{fig:fishing_raster_diff}C). The spatial redistribution patterns of displaced vessels relative to non-displaced vessels suggest that some relocated to other waters in Kiribati (\emph{i.e.} Gilbert islands and Line islands), but also the Marshall Islands, Tuvalu, Nauru, and the high seas.

@@

We obtained information on revenues from the Pacific Islands Forum Fisheries Agency \emph{Tuna Development Indicators 2016} report.  Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA \cite{ffa_2017}) where annual revenues from license fees (for VDS and other access programs) are reported for each country (2008 - 2016; Fig. \ref{fig:empirical}C; Fig. S8 - S9). For countries in the PNA, these revenues show a combination of vessel-day license fees as well as joint-venture operations.

Shapefiles of Exclusive Economic Zones were obtained through Marine Ecoregions of The World, we use World EEZ v10 (2018-02-21) available for download at: http://www.marineregions.org. Shapefiles for Marine Protected Areas come from the World Database of Protected Areas, and were downloaded in March 2019 from: https://www.protectedplanet.net. All analyses were performed in R version 3.6.1 and RStudio 1.2.5.001 \cite{rcore_2019}.

@@

Juan Carlos Villaseñor-Derbez was supported by UCMexus-CONACyT (CVU: 669403) and the Latin American Fisheries Fellowship Program. John Lynham was supported by the Conservation Strategy Fund, with a grant from the Pew Charitable Trusts and Pew Bertarelli Ocean Legacy. John Lynham and Christopher Costello wish to acknowledge The Nature Conservancy.

@@

% Map of LSMPAS in PNA
\begin{figure}
\centering
\includegraphics{img/PNA_map.png}
\caption{\label{fig:PNA_map}Map of the Exclusive Economic Zones (EEZs) and Marine Protected Areas in the PNA, the bottom-left inset presents provides a reference for the area. Parties to the Nauru Agreement (PNA) are shown in blue, while empty polygons indicate all others. A red line indicates the Kiribati EEZ. The solid red polygons show The Phoenix Islands Protected Area implemented in 2015 and the proposed Palau National Marine Sanctuary to be implemented in 2020. Land masses are shown in gray. Labels indicate ISO3 country codes for PNA members (PLW: Palau, PNG: Papua New Guinea, FSM: Federal States of Micronesia, SLB: Solomon Islands, NRU: Nauru, MHL: Marshal Islands, KIR: Kiribati, TUV: Tuvalu, TKL: Tokelau).}
\end{figure}

% PNA model output for costs
\begin{figure}[htbp]
\centering
\includegraphics{img/PNA_model.pdf}
\caption{\label{fig:PNA_model}Cost of spatial closures in a vessel-day fishery. Each line represents a possible value of within-country stock movement ($\theta$; line colors), with $\theta = 0$ representing a stock that doesn't move to $\theta = 1$ depicting a stock that continuously moves between the reserve and the fishing zone. The revenue losses to Country 1 (vertical axis) are relative to a fishery with no spatial closures, and are shown as a function of reserve size ($R$; horizontal axis), from no reserve ($R = 0$) to closing the entire Exclusive Economic Zone ($R = 1$). Costs are shown for Country 1 when there is no trading (A) and when trading is allowed (B; note the change in axis limits). Costs avoided by trading are shown in (C). Dashed red line in (A) is a 1:1 line. When trading between countries occurs, 88\% to 99\% of revenue losses can be avoided.}
\end{figure}
